Exploring Constraint Grammar as a SAT problem

### 2015-03-26 paper/nodalida2015.tex
removed some text, added some TODOs

\paragraph{Rules are unordered.}
A typical CG rule sequence starts with rules that are either very common, or have a complex condition to fulfil.
Stronger rules appear towards the end, to be applied if none of the weaker rules has fired. 
For a rule set such as following, this order is crucial:

@@

Finite-state implementations of CG \cite{koskenniemi92} \footnote{\todo{Inari: check Tapanainen 1996 if there's anything about ordering}} also have logically unordered rules.

To make up for the lack of sequential order, we have experimented with different strategies of maximising and solving conflicts:
\begin{itemize}
\item Number of rules used
\item Number of instances when a rule is applied
\item 
\end{itemize}

@@

SAT is NP-complete, whereas the current implementations of CG are polynomial, 
but with advances in SAT solving techniques, the performance isn't as catastrophically bad as in 90s-00s.
We used the open-source SAT solver MiniSat \cite{een04sat}.
We do foreach sentence: foreach rule; 
%whereas VISL CG3 has foreach rule: foreach sentence.
the size of the SAT problem is dependent of the sentence (???).

The table below shows
Note also that this is a first proof-of-concept implementation and could be optimised.

\begin{table}
  \centering
  \begin{tabular}{|l|c|c|}
     \hline
	         & Quijote      & Pride \\ \hline
  VISL CG3 & 0m10.745s & 0m4.811s \\ \hline
  SATCG    & 2m53.708s & 0m55.705s \\ \hline
  \end{tabular}
  \caption{Comparing time between SATCG and VISL CG3.}
  \label{table:time}
\end{table}

@@

The tendency of needing less rules to disambiguate shows also in a larger context with more complex rules.
SATCG disambiguated more 17 \% of the cases for Pride and Prejudice, and 28 \% of the cases for Don Quijote.

@@

In this experiment, we test the performance of the two implementations by working with a smaller grammar and comparing the result to

>>>>

\paragraph{Rules are unordered.}
A typical CG rule sequence starts with rules that are either very common, or have a complex condition to fulfil.
Stronger rules appear towards the end, to be applied if none of the weaker rules has. 
For a rule set such as following, this order is crucial:

@@

To replace the sequential order, we have experimented with different strategies:
\begin{itemize}
\item Maximise number of rules used---assume if someone has written a rule, it is worth applying. Didn't give the best results.
\item Maximise number of instances when a rule is applied---\todo{comment about this too}
\end{itemize}

% Finite-state implementations of CG \cite{koskenniemi92} also have logically unordered rules. 
% \todo{Inari: check if Tapanainen 1996 discuss ordering}

These three features are likely to change the way that rules are written. \todo{more speculation, examples}

@@

SAT is NP-complete, whereas the current implementations of CG are polynomial, 
but with advances in SAT solving techniques, 
the performance is much more feasible than in the previous works done in 90s--00s.
We used the open-source SAT solver MiniSat \cite{een04sat}.
%We do foreach sentence: foreach rule; 
%whereas VISL CG3 has foreach rule: foreach sentence.
%the size of the SAT problem is dependent of the sentence (???)

Table~\ref{table:time} shows the execution time compared to VISL CG3.
SATCG is just a first proof-of-concept implementation and can be optimised.

\begin{table}
  \centering
  \begin{tabular}{|l|c|c|}
     \hline
	         & \textbf{Don Quijote}  & \textbf{Pride and Prejudice} \\ \hline
  VISL CG3 & 0m10.745s & 0m4.811s \\ \hline
  SATCG    & 2m53.708s & 0m55.705s \\ \hline
  \end{tabular}
  \caption{Comparing time between SATCG and VISL CG3.}
  \label{table:time}
\end{table}

@@

The tendency of needing less rules to disambiguate shows also in a larger context with more complex rules. 
SATCG disambiguated more 17 \% of the cases for Pride and Prejudice, and 28 \% of the cases for Don Quijote. \todo{Analysis of those cases}

@@

In this experiment, we test the performance of the two implementations by working with a smaller grammar. We compare how much removing the rules affects the result, gold standard being for each implementation the result given by the full rule set.
\todo{Perform this experiment}
